---
title: "RWorksheet_Subosa#4c.Rmd"
author: "Gian Adree Subosa"
date: "2024-11-09"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1. Use dataset mpg
```{r}
#a.) Show your solutions on how to import a csv file into the environment.
mpg_file <- read.csv("mpg.csv")
head(mpg_file)
str(mpg_file)
```

```{r}
#b.) Which variables from mpg dataset are categorical?

#According to my observation, the variables from the mpg dataset that are seemly categoral are the following:
# - manufacturer: that indicates the manufacturer name
# - model: that indicates the model name
# - trans: the type of transmission system the vehicle uses 
# - drv: indicating whether the vehicle is front-wheel drive (f), rear-wheel drive (r), or four-wheel drive (4)
# - fl: the type of fuel
# - class: the general category or type of the vehicle
```

```{r}
#c.) Which are continuous variables?

#The following are the continuous variables in the mpg dataset:
# - dspl: the engine displacement in liters
# - year: the year of manufacture
# - cyl: the number of cylinders
# - cty: the city miles per hour
# - hwy: the highway miles per gallon
```

#2.1 Which manufacturer has the most models in this data set? Which model has the most variations?
```{r}
#a.) Group the manufacturers and find the unique models
library(dplyr)
data(mpg, package = "ggplot2")

modelmanufacturer <- mpg %>%
  group_by(manufacturer) %>%
  summarize(unique_models = n_distinct(model)) %>%
  arrange(desc(unique_models))

print(modelmanufacturer)

modelvariations <- mpg %>%
  group_by(model) %>%
  summarize(variations = n()) %>%
  arrange(desc(variations))

print(modelvariations)
```

```{r}
#b.) Graph the result by using plot() and ggplot()
modelmanufacturer$manufacturer <- factor(modelmanufacturer$manufacturer, levels = modelmanufacturer$manufacturer)

barplot(modelmanufacturer$unique_models, 
        names.arg = modelmanufacturer$manufacturer,
        main = "The Number of Unique Models by Manufacturer",
        xlab = "Manufacturer",
        ylab = "Unique Models",
        col = "pink",
        las = 2, 
        cex.names = 0.7)
```

#2.2 Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.
```{r}
#a.) What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
library(ggplot2)
ggplot(mpg, aes(x = model, y = manufacturer)) + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Relationship of Car Models per Manufacturer", 
       x = "Car Model",
       y = "Manufacturer"
  )
#The data is presented in the form of a scatter plot, with a point at each model and manufacturer combination. It depicts which manufacturers produce which models, but it does not properly represent the number of variations for each model because it just displays the presence or absence of each combination.
```

```{r}
#b.) For you, is it useful? If not, how could you modify the data to make it more informative?

#Because of the overlapping points, this visualization is currently ineffective. A better approach could be summarizing and color-code by count that display the counts of differences in size or color intensity and/or consider inverting the coordinates for better reading. Here's how you can modify the data:

modelmanufacturer <- mpg %>%
  group_by(model, manufacturer) %>%
  summarize(variations = n())

ggplot(modelmanufacturer, aes(x = model, y = manufacturer, size = variations)) +
  geom_point(color = "red", alpha = 0.6) +
  labs(title = "Relationship of Model and Manufacturer Variations", x = "Model", y = "Manufacturer") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

#3. Plot the model and the year using ggplot(). Use only the top 20 observations.
```{r}
library(ggplot2)
top_20_mpg <- head(mpg, 20)

ggplot(top_20_mpg, aes(x = model, y = year)) +
  geom_point(color = "green", size = 3) +
  labs(title = "Top 20 Observations for Model by Year", x = "Model", y = "Year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#4. Using the pipe (%>%), group the model and get the number of cars per model.
```{r}
library(dplyr)

cars_p_model <- mpg %>%
  group_by(model) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

print(cars_p_model)
```

```{r}
#a.) Plot using geom_bar() using the top 20 observations only.
library(ggplot2)
library(dplyr)

mpg_top_20 <- mpg %>%
  arrange(desc(year))
  head(20)
  
ggplot(mpg_top_20, aes(x = reorder(model, -year), fill = manufacturer)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Top 20 Car Models Per Year", x = "Car Model", y = "Number of Cars") +
  scale_fill_brewer(palette = "Set3")
```

```{r}
#b.) Plot using the geom_bar() + coord_flip() just like what is shown below
library(ggplot2)
library(dplyr)

mpg_top_20 <- mpg %>%
  arrange(desc(year)) %>%  
  head(20)


ggplot(mpg_top_20, aes(x = reorder(model, -year), fill = manufacturer)) + 
  geom_bar() +  
  coord_flip() +  
  labs(title = "Top 20 Car Models Per Year", x = "Count of Cars", y = "Car Model") +
  scale_fill_brewer(palette = "Set2") +  
  theme_minimal() + 
  theme(legend.title = element_blank())  
```

#5. Plotting the relationship between cyl and displ
```{r}
#a.) How would you describe its relationship? Show the codes and its result.
library(ggplot2)

ggplot(mpg, aes(x = cyl, y = displ, color = displ)) + 
  geom_point(size = 3) +  
  labs(title = "Relationship between No. of Cylinders and Engine Displacement", 
       x = "Number of Cylinders", 
       y = "Engine Displacement") +
  scale_color_gradient(low = "blue", high = "red") +  
  theme_minimal() 

#The scatter plot indicates that engine displacement (displ) and cylinder count (cyl) are positively correlated. Engine displacement often rises in tandem with the number of cylinders. In line with automotive design, where greater displacement usually translates into higher power and performance, this pattern implies that cars with more cylinders typically have larger engines.
```

#6.1 Plot the relationship between displ (engine displacement) and hwy(highway miles per gallon). Mapped it with a continuous variable you have identified in #1-c. What is its result? Why it produced such output?
```{r}
library(ggplot2)

ggplot(mpg, aes(x = displ, y = hwy, color = displ)) + 
  geom_point(size = 3) +  
  labs(title = "Relationship between Engine Displacement and Highway MPG", 
       x = "Engine Displacement (L)", 
       y = "Highway MPG") +
  scale_color_gradient(low = "blue", high = "red") + 
  theme_minimal()  

#This scatter plot depicts the relationship between displ and hwy, with color indicating the number of cylinders (cyl). Cars with larger engine displacement (displ) typically have fewer highway miles per gallon (hwy), indicating a trade-off between engine power and fuel efficiency. The output most likely reflects that vehicles with bigger engine displacement and more cylinders consume more fuel, resulting in lower highway numbers.
```

#6.2 Import the traffic.csv onto your R environment.
```{r}
#a.) How many numbers of observation does it have? What are the variables of the traffic dataset?
traffic_file <- read.csv("traffic.csv")

num_observations <- nrow(traffic_file)
num_observations

variable <- colnames(traffic_file)
variable

#The data has a total of 48120 observations and its variables are "DateTime", "Junction", "Vehicles", and "ID".
```

```{r}
#b.) Subset the traffic dataset into junctions.
traffic_data <- read.csv("traffic.csv")
junctions_list <- split(traffic_data, traffic_data$Junction)
junctions_list[[1]]
```

```{r}
#c.) Plot each junction in a using geom_line()
library(ggplot2)
traffic_file <- read.csv("traffic.csv")
traffic_file$DateTime <- as.POSIXct(traffic_file$DateTime, format="%Y-%m-%d %H:%M:%S")
ggplot(traffic_file, aes(x = DateTime, y = Vehicles)) +
  geom_line() +
  facet_wrap(~ Junction) +
  labs(title = "Vehicle Counts at Each Junction Over Time", x = "DateTime", y = "Vehicles") +
  theme_minimal()
```

#7. From alexa_file.xlsx, import it to your environment
```{r}
#a.) How many observations does alexa_file has? What about the number of columns?
library(readxl)
alexa_file <- read_excel("alexa_file.xlsx")
num_obs <- nrow(alexa_file)
num_cols <- ncol(alexa_file)

print(num_obs)
print(num_cols)

#The data shows that "alexa_file.xlsx" has a total of 3150 obsevations, and there are 5 number of columns.
```

```{r}
#b.) Group the variations and get the total of each variations. Use dplyr package.
library(dplyr)
library(readxl)
alexa_file <- read_excel("alexa_file.xlsx")
total_variation <- alexa_file %>%
  group_by(variation) %>%
  summarise(total = n())

print(total_variation)
```

```{r}
#c.) Plot the variations using the ggplot() function. What did you observe?
library(ggplot2)
library(dplyr)
library(readxl)
alexa_file <- read_excel("alexa_file.xlsx")
total_variation <- alexa_file %>%
  group_by(variation) %>%
  summarise(total = n())

ggplot(total_variation, aes(x = variation, y = total, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Number of Each Variation", x = "Variation", y = "Total Number") +
  theme_minimal()

#This bar chart shows which variation/s are most frequently reviewed, providing insight into consumer preferences or distribution trends for each variation. And I noticed that the color of the bar plot is in rainbow since I didn't use the "col" variable in my code.
```

```{r}
#d.) Plot a geom_line() with the date and the number of verified reviews.
review_counts <- alexa_file %>%
  group_by(date) %>%
  summarise(verified_reviews = sum(feedback))

ggplot(review_counts, aes(x = date, y = verified_reviews)) +
  geom_line(color = "orange") +
  labs(
    title = "Verified Reviews Chart",
    x = "Date",
    y = "Number of Verified Reviews"
  ) +
  theme_minimal()
```

```{r}
#e.) Get the relationship of variations and ratings. Which variations got the most highest in rating? Plot a graph to show its relationship.
library(dplyr)
library(ggplot2)
library(readxl)
alexa_file <- read_excel("alexa_file.xlsx")
variation_ratings <- alexa_file %>%
  group_by(variation) %>%
  summarise(avg_rating = mean(rating, na.rm = TRUE))

ggplot(variation_ratings, aes(x = variation, y = avg_rating, fill = variation)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Rating per Variation", x = "Variation", y = "Average Rating") +
  theme_minimal()

highest_rating_variation <- variation_ratings %>%
  filter(avg_rating == max(avg_rating))

print(highest_rating_variation)

#According to the result, the "Walnut Finish" is the highest rating variation with 4.8 average rating.
```

