---
title: "RWorksheet_Subosa"
author: "Gian Adree Subosa"
date: "2024-09-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1.Set up a vector named age
``` {r setup}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41)

num_data_points <-length(age)

#How many data points?
num_data_points
```
#2.Reciprocal age
``` {r setup}
#Write the code...
library(MASS)
reciprocal_age <- fractions(1 / age)
#...and its output
reciprocal_age
```
#3. Assign
``` {r setup}
new_age <- c(age, 0, age)

#What happens to new_age?
new_age
```
#4. Sort
```{r setup}
#Write the code and its output
sort(age)
```
#5. Find minimum and maximum value for age
``` {r setup}
#These are the codes:
min_age <- min(age)
max_age <- max(age)

#These are the outputs:
min_age
max_age
```
#6. Set up a vector named data
```{r setup}
data <- c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)

#How many data points? Write the R code and its output.
num_data_points <- length(data)
num_data_points
```
#7. Generate a new vector for data where you double every value of the data
```{r setup}
new_data <- c(data ^ 2)
new_data
#It was noticed that the value of data itself has been doubled after using the "^" operator.
```
#8.1 and 8.2 Generate a sequence
```{r setup}
#Integers from 1 to 100
seq_8.1 <- 1:100
seq_8.1
#Numbers from 20 to 60
seq_8.2 <- 20:60
seq_8.2
```
#8.3 Mean of numbers from 20 to 60
```{r setup}
mean_numbers <- mean(seq_8.2)
mean_numbers
```
#8.4 Sum of numbers from 51 to 91
```{r setup}
sum_numbers <- sum(51:91)
sum_numbers
```
#8.5 Integers from 1 to 1000
```{r setup}
seq_8.5 <- 1:1000
seq_8.5

#Find the maximum  value until 10.
sequence_8.5 <- 1:10
sequence_8.5
length_8.5 <- length(sequence_8.5)
length_8.5
```
#9 Filter Option
```{r setup}
#Write the code...
integers <- seq(100)
filtered_integers <- Filter(function(i) { all(i %% c(3,5,7) != 0) }, seq(100))
#...and its output
filtered_integers
```
#10 Generate a sequence backwards of the integers from 1 to 100.
```{r setup}
#Write the code...
sequence_i <- 1:100
rev_sequence_i <- rev(sequence_i)
#...and its output
print(rev_sequence_i)
```
#11. List all the natural numbers below 25 that are multiples of 3 or 5.
```{r setup}
#Natural numbers
numbers <- 1:24
#Find numbers that are multiples of 3 or 5
multiples_of_3_or_5 <- numbers[numbers %% 3 == 0 | numbers %% 5 == 0]
#Sum of multiples
sum_multiples <- sum(multiples_of_3_or_5)
#The outputs are:
multiples_of_3_or_5
sum_multiples
```
#12 Statements.
```{r setup}
x <- {0 + x + 5 + }

# The vector indicates that R encountered an unexpected end of input because the expression is not complete. R is expecting more input to finish the statement.
```
#13. Set up a vector named score
```{r}
#Set up the vector
score <- c(72, 86, 92, 63, 88, 89, 91, 92, 75, 75, 77)
x2 <- score[2]
x3 <- score[3]
#The outputs are
x2
x3
```
#14 Create another vector.
```{r}
# Create the vector
a <- c(1, 2, NA, 4, NA, 6, 7)
#Print the vector with NA values displayed as -999
print(a, na.print = "-999")
#output 
#     2 -999    4 -999    6    7 
#Original Vector: a is c(1, 2, NA, 4, NA, 6, 7). Printing with na.print: NA values are displayed as -999 in the output. The actual vector a stays the same; only how it’s printed is affected.
```
#15 Special function.
```{r setup}
#Prompt for name and age.
name = readline(prompt="Input your name: ")
age = readline(prompt="Input your age: ")
#Print the output.
print(paste("My name is", name, "and I am", age, "years old."))
```